Fail Faster

### 2025-03-13 paper/paper.tex
CP 4.3

\subsection{Let-Insertion and Effect Ordering}
Careful readers might note that the definition of \texttt{bind} in Section~\ref{subsection:basic-design}
was more complicated than one might expect. In particular, why not define bind in a more standard way: as
\texttt{let bind' g k = fun size random -> k (g size random) size random}, without the code block
that let-binds the spliced code \texttt{.\~{}(g size random)}?
Unfortunately, using \texttt{Gen.bind'} leads to incorrect code being generated.
For example, consider \texttt{Gen.bind' (Gen.int .<0>. .<1>.) (fun x -> Gen.return .<(.~x,.~x)>.)}.
This generates the run time code \texttt{fun size random -> (SR.int random 0 1, int SR.random 0 1)},
which is incorrect. 
This is not equivalent to the behavior of writing the same code with base quickcheck \jwc{Which is a property we want to hold for our eval.}.
Instead of generating a single integer and returning it twice, it samples two different integers.

@@

The library \jwc{ensure this is consistent with the way we talk about the project, cf cross-language} is careully designed to 
preserve exactly the effect order of base quickcheck \jwc{and the scala version to preserve the effect ordering of ScalaCheck}.
\jwc{This fact should go earlier. Not really sure about where this subsection should actually land, but it needs to be said somehwere.}

>>>>

\subsection{Let-Insertion and Effect Ordering}
Careful readers might note that the definition of \texttt{bind} in Section~\ref{subsection:basic-design}
was more complicated than one might expect. In particular, why not define bind in a more standard way: as
\texttt{let bind' g k = fun size random -> k (g size random) size random}, without the code block
that let-binds the spliced code \texttt{.\~{}(g size random)}?\hg{Need to spell
this out --- the vast majority of readers won't have that precise question in
their heads. Might even be worth comparing the two implementations here side by side}
Unfortunately, using \texttt{Gen.bind'} leads to incorrect code being generated.
For example, consider \texttt{Gen.bind' (Gen.int .<0>. .<1>.) (fun x -> Gen.return .<(.~x,.~x)>.)}.
This generates the run time code \texttt{fun size random -> (SR.int random 0 1, int SR.random 0 1)},
which is incorrect. 
This is not equivalent to the behavior of writing the same code with base quickcheck \jwc{Which is a property we want to hold for our eval.}.
Instead of generating a single integer and returning it twice, it samples two different integers.

@@

The library \jwc{ensure this is consistent with the way we talk about the project, cf cross-language} is careully designed to 
preserve exactly the effect order of base quickcheck \jwc{and the scala version to preserve the effect ordering of ScalaCheck}.
\jwc{This fact should go earlier. Not really sure about where this subsection should actually land, but it needs to be said somehwere.}
\hg{+1, I think this feels sort of out of place here, but I also think moving it
up would make that part harder to understand too. Is there a way to tie this in
to our discussion about maintaining the precise set of choices? The two issues
are different, but they're related}
